### 2017-07-17 lafaySimulatedScenePerception.tex

The previous experiment demonstrated that, among the classes of sounds occurring in urban soundscapes, those with markers are specific to some of environments. Those sound markers seem to have a great impact on perception. This impact is studied here in more detail using an added benefit of the simulation paradigm proposed in this paper, the ability to manipulate and modify the generated scenes.

@@

In order to verify that the pleasantness of the ideal and non ideal scenes if affected by the sound markers, the audio waveform of the simulated scenes are regenerated with or without considering the classes identified as markers. For the i-scenes, some of the i-markers are removed. For the ni-scenes, some of the ni-markers are removed. A perceptive test evaluating the pleasantness is then conducted with a protocol close to the one considered in experiment 1.b.

@@

Thus the objective of this experiment is to study if the removal of the previously identified markers have an impact on the perceived pleasantness. Two hypothesis are thus formulated:

@@

If the first hypothesis is rather intuitive, the second is less. Indeed, it is not obvious that the removal of the i-markers, though perceptively connoted, decreases the global quality of a soundscape. This removal also decreases the global level of the scene.

@@

\begin{itemize}
\item \emph{72 km-scenes}: the 72 scenes originally simulated by the subject of experiment 1.a where the sound classes identified as markers are kept. The 36 ideal scenes with markers are noted i/km-scenes, and the 36 non ideal scenes with markers are noted ni/km-scenes.
\item \emph{72 rm-scenes}:  72 scenes where the sound classes identified as markers are removed. The 36 ideal scenes without markers are noted i/rm-scenes, and the 36 non ideal scenes without markers are noted ni/rm-scenes.
\end{itemize}

@@

In order to create km-scenes with still some sound diversity and no long duration absence of sound activity within the scene, only the sound classes of events of the first level of abstraction are removed, see Table~\ref{tab:markers}. Those classes are:

@@

For each subject, the scenes are presented in a random order. The first 10 scenes allow the subject to calibrate their scores. Those calibration scenes are  5 i/km-scenes and 5 ni/km-scenes. These first 10 scenes are replayed at the end of the experiment, and only the notes given at the second occurrence are taken into account.

@@

All the subjects perform the test on the same type of computers . Thy listen to the stimuli with semi-open \emph{Beyer-Dynamic DT 990 Pro} headphones. The sound level is the same for all subjects.

@@

A supervisor is available during the whole duration of the experiment to ensure a smooth running of the latter and answer any questions the subjects may have.

@@

The aim here is to validate the hypothesis that the removal of i-markers and ni-markers have an impact on the perceived pleasantness. To do so, we perform a variance analysis. We consider $\mathcal{A}_{sujet}$ as a dependent variable, and as independent variables, the type of environment (i/ni), and the presence/absence of markers (km/rm). As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to check the validity of the hypothesis that there exist a significant difference of perceived pleasantness between km-scenes and rm-scenes. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: i/ni; marker: km/rm), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analysis are done by following the Tukey-Kramer procedure.

@@

Let us consider $\mathcal{A}_{sujet}$ for the km-scenes. Close inspection that the answers given by one subject strongly differs with the ones of others. This subject evaluated positively close to half of the ni/km-scenes, see Annex~\ref{app:xp2}, subject 7. This subject also gave a score above 0 for 58\% of the ni/km-scenes, contrary to an average of 11\% for the remaining of the subjects.

Furthermore, this subject used the whole scale (from -5 to 5) to score both the i-scenes and the ni-scenes. Those behaviors strongly differing with the ones of the others subjects be they from this experiment or the previous ones, subject 7 is discarded from the analysis.

@@

In this section, we study the scores given by the subject while listening to the several types of scenes, namely the i/km-scenes, ni/km-scenes, i/rm-scenes et ni/rm-scenes. The repeated measure ANOVA applied to $\mathcal{A}_{sujet}$ show a significative effect of the type environment (i/ni: $F[1,10]=175$, $p<0.01$), of the presence/absence of markers (km/rm: $F[1,10]=7$, $p<0.05$), and of the interaction between those two factors ($F[1,10]=67$, $p<0.01$).

@@

The \emph{post hoc} analysis exhibit significative differences between all the groups of observations, notably between the i/km-scenes ($p<0.05$) and the i/rm-scenes ($p<0.01$).

@@

The i-markers do have a positive impact on the perception of an environment. The fact that their removal decreases $\mathcal{A}_{scene}$ clearly shows that it is possible to improve the perceived pleasantnes of a given urban area by the addition of sounds commonly considered as pleasant such as \emph{bird calls}. Those conclusions are in line with the positive approacj introduced by Schafer in \cite{schafer1977tuning}.

@@

These results tend to confirm that the perception of the qualities of a scene indeed depends above all on the sound sources that compose it. The compositional characteristics that are taken into account during the perceptual process appear to vary from one source to the other, from one type of environment to the other. This fact lead the authors to conclude that there is little hope to find a holistic physical descriptor that can adequately account for the affective qualities of all types of sound environment.

@@

\begin{itemize}
\item in the case of non ideal scenes, one should focus on reducing the acoustic pressure level, whether globally, or by discarding specific sources such as \emph{sirens} or \emph{car horns}
\item in the case of ideal scenes, one should first identity which sources are pleasant to the targeted community, second lower the volume of the other sound sources, and, if possible, raise the contribution or add positive sound markers.
\end{itemize}

@@

The outcomes of the three experiments described in this paper demonstrate the usefulness of considering simulation tools such as \emph{simScene} in order to scientifically question the perception of soundscapes. We also believe that its wider usage could enable urban planning decision-makers to question an entire community about its own representations of the sound environment to which it is exposed, and about the representations of the sound environments to which it would like to be exposed.

@@

Future work should consider several other simulation experiments by varying the emotional qualities (quiet, comfortable, troublesome, etc.), but also by specifying particular places (park, square, street, etc.), in order to provide to the scientific community entire corpora of cognitively informed soundscapes.

@@

Finally, one should study the influence of socio-cultural contexts on perception. Indeed, if the sound of the church bell is most often a quality environment marker for a Westerner, this is not necessarily true with subjects of Eastern, Middle Eastern or other cultures.

@@

\begin{itemize}
\item The simulator can be deployed on a large scale via the Internet thanks to the chosen software architecture;
\item simulated scenes can be analyzed without having to take into account the different mother tongues of the subjects, the semantic nature of the classes of sounds used being known \emph{a priori} by the experimenter and without the need to analyze and annotate the sound scene to identify the sources, their occurrences in the simulated scene being directly available.
\end{itemize}

>>>>

The previous experiment demonstrated that, among the classes of sounds occurring in urban soundscapes, those with markers are specific to some environments. Those sound markers seem to have a great impact on perception. This impact is studied here in more detail using an added benefit of the simulation paradigm proposed in this paper, the ability to manipulate and modify the generated scenes.

@@

In order to verify that the pleasantness of the ideal and non ideal scenes is affected by the sound markers, the audio waveform of the simulated scenes are regenerated with or without using the classes identified as markers. For the i-scenes, some of the i-markers are removed. For the ni-scenes, some of the ni-markers are removed. To evaluate the impact on the perception of pleasantness caused by those modifications, a perceptive test is then conducted with a protocol close to the one considered in experiment 1.b.

@@

The objective of this experiment is to study if the removal of the previously identified markers have an impact on the perceived pleasantness. Two hypothesis are thus formulated:

@@

If the first hypothesis is rather intuitive, the second is less. Indeed, it is not obvious that the removal of the i-markers, though perceptively positively connoted, decreases the global quality of a soundscape. Indeed, this removal also decreases the global level of the scene.

@@

\begin{itemize}
\item \emph{72 km-scenes}: the 72 scenes originally simulated by the subject of experiment 1.a where the sound classes identified as markers are kept. The 36 ideal scenes with markers are noted i/km-scenes, and the 36 non ideal scenes with markers are noted ni/km-scenes (k for kept).
\item \emph{72 rm-scenes}:  72 scenes where the sound classes identified as markers are removed. The 36 ideal scenes without markers are noted i/rm-scenes, and the 36 non ideal scenes without markers are noted ni/rm-scenes (r for removed).
\end{itemize}

@@

In order to create km-scenes with still some sound diversity and no  absence of sound activity within a long period of the scene, only the sound classes of events of the first level of abstraction are removed, see Table~\ref{tab:markers}. Those classes are:

@@

For each subject, the scenes are presented in a random order. The first 10 scenes allow the subject to calibrate their scores. Those calibration scenes are  5 i/km-scenes and 5 ni/km-scenes. These first 10 scenes are replayed at the end of the experiment, and only the scores given at the second occurrence are taken into account.

@@

All the subjects perform the test on the same type of computers. They listen to the stimuli with semi-open \emph{Beyer-Dynamic DT 990 Pro} headphones. The sound level is the same for all subjects.

@@

A supervisor is available during the whole duration of the experiment to monitor the experiment and answer any questions the subjects may have.

@@

The aim here is to validate the hypothesis that the removal of i-markers and ni-markers have an impact on the perceived pleasantness. To do so, we perform a variance analysis. We consider $\mathcal{A}_{sujet}$ as a dependent variable, and as independent variables, the type of environment (i/ni), and the presence/absence of markers (km/rm). As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to check the validity of the hypothesis that there exist a significant difference of perceived pleasantness between km-scenes and rm-scenes. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: i/ni,  marker: km/rm), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analyses are done using the Tukey-Kramer procedure.

@@

Let us consider $\mathcal{A}_{sujet}$ for the km-scenes. Close inspection of the answers show that one subject strongly differs with the ones of others. This subject evaluated positively close to half of the ni/km-scenes, see Annex~\ref{app:xp2}, subject 7. This subject also gave a score above 0 for 58\% of the ni/km-scenes, contrary to an average of 11\% for the remaining of the subjects.

Furthermore, this subject used the whole scale (from -5 to 5) to score both the i-scenes and the ni-scenes. Those behaviors strongly differ with the ones of the others subjects be they from this experiment or the previous ones. Subject 7 is thus discarded from the analysis.

@@

In this section, we study the scores given by the subject while listening to the several types of scenes, namely the i/km-scenes, ni/km-scenes, i/rm-scenes and ni/rm-scenes. The repeated measure ANOVA applied to $\mathcal{A}_{sujet}$ show a significative effect of the type environment (i/ni: $F[1,10]=175$, $p<0.01$), of the presence/absence of markers (km/rm: $F[1,10]=7$, $p<0.05$), and of the interaction between those two factors ($F[1,10]=67$, $p<0.01$).

@@

The \emph{post hoc} analysis exhibits significative differences between all the groups of observations, notably between the i/km-scenes ($p<0.05$) and the i/rm-scenes ($p<0.01$).

@@

The i-markers do have a positive impact on the perception of an environment. The fact that their removal decreases $\mathcal{A}_{scene}$ clearly shows that it is possible to improve the perceived pleasantnes of a given urban area by the addition of sounds commonly considered as pleasant such as \emph{bird calls}. Those conclusions are in line with the positive approach introduced by Schafer in \cite{schafer1977tuning}.

@@

These results tend to confirm that the perception of the qualities of a scene indeed depends above all on the sound sources that compose it. The compositional characteristics that are taken into account during the perceptual process appear to vary from one source to the other, from one type of environment to the other. This fact lead the authors to believe that there is little hope to find a holistic physical descriptor that can adequately account for the affective qualities of all types of sound environment.

@@

\begin{itemize}
\item in the case of non ideal scenes, one should focus on reducing the acoustic pressure level, whether globally, or by discarding specific sources such as \emph{sirens} or \emph{car horns}
\item in the case of ideal scenes, one should first identify which sources are pleasant to the targeted community, second lower the volume of the other sound sources, and, if possible, raise the contribution or add positive sound markers.
\end{itemize}

@@

The outcomes of the three experiments described in this paper demonstrate the usefulness of considering a dedicated simulation tool such as \emph{simScene} in order to scientifically question the perception of soundscapes in an innovative way. We also believe that its wider usage could enable urban planning decision-makers to question an entire community about its own representations of the sound environment to which they are exposed, and about the representations of the sound environments to which they would like to be exposed.

@@

Future work should consider several other simulation experiments by varying the emotional qualities (quiet, comfortable, troublesome, etc.), but also by specifying specific urban places (park, square, street, etc.), in order to provide to the scientific community an entire corpora of cognitively informed soundscapes.

@@

Finally, one should study the influence of socio-cultural contexts on perception. Indeed, if the sound of the church bell is most often a quality environment marker for a Westerner, this do not necessarily hold true for subjects of Eastern, Middle Eastern or other cultures.

@@

\begin{itemize}
\item The simulator can be deployed on a large scale via the Internet thanks to the chosen software architecture;
\item simulated scenes can be analyzed without the need to take into account the different mother tongues of the subjects, the semantic nature of the classes of sounds used being known \emph{a priori} by the experimenter and without the need to analyze and annotate the sound scene to identify the sources, their occurrences in the simulated scene being directly available.
\end{itemize}
